Acoustic Scene Classification: \\ Supervised and Unsupervised Approaches using the Scattering Transform

### 2016-06-22 paper/structureScene.tex
minor improvements to section 3

\subsection{Temporal scattering}
Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
dimensionless frequency $1$ and bandwidth $1/Q_1$.
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
according to a geometric sequence of temporal scales $2^{-\gamma_1/Q_1}$.
The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
In the sequel, the number of filters per octave $Q_1$ is set to $8$, and
the number of octaves $J_1$ is set to $10$, \ie the full audible range.
We denote by
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(-2^{\gamma_1/Q_1} t)
\end{equation}
the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
has a center frequency of $2^{-\gamma_1/Q_1}$, a bandwidth of $2^{-\gamma_1/Q_1}/Q_1$, and a quality factor of $Q_1$.

The wavelet transform $\boldsymbol{}$ of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
$\boldsymbol{y_1}(t, \gamma_1) =
(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{y_1}(t, \gamma_1)\vert
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).
\end{equation}
The wavelet scalogram bears resemblance with the constant-Q transform (CQT),
which is derived from the short-time Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the CQT has a fixed temporal resolution, constrained by the support of the STFT analyzing window.
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies, which allows to observe steady amplitude modulations.

In all subsequent experiments,
$\boldsymbol{\psi}$ is designed as a fourth-order Gammatone
wavelet of quality factor $Q = 8$.

>>>>

\subsection{Temporal scattering}
Let $\boldsymbol{\psi}(t)$ a complex-valued band-pass filter of
dimensionless frequency $1$ and bandwidth $1/Q_1$.
A filter bank of wavelets is built by dilating $\boldsymbol{\psi}(t)$
according to a geometric sequence of temporal scales $2^{-\gamma_1/Q_1}$.
The variable $\gamma_1$, akin to a log-frequency, takes integer values between $0$ and $(J_1 \times Q_1 - 1)$.
In the sequel, the number of filters per octave $Q_1$ is set to $8$, and
the number of octaves $J_1$ is set to $10$, \ie the maximum range of human hearing.
We denote by
\begin{equation}
\boldsymbol{\psi_{\gamma_1}}(t) = 2^{-\gamma_1/Q_1} \boldsymbol{\psi}(-2^{\gamma_1/Q_1} t)
\end{equation}
the wavelets resulting from the dilation of $\boldsymbol{\psi}$(t).
For some $\gamma_1$, the wavelet $\boldsymbol{\psi_{\gamma_1}}(t)$
has a center frequency of $2^{-\gamma_1/Q_1}$, a bandwidth of $2^{-\gamma_1/Q_1}/Q_1$, and a quality factor of $Q_1$.

The wavelet transform $\boldsymbol{}$ of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelets:
$\boldsymbol{y_1}(t, \gamma_1) =
(\boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}})(t)$.
Applying pointwise complex modulus to $\boldsymbol{y_1}$ yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{y_1}(t, \gamma_1)\vert
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t).
\end{equation}
The wavelet scalogram bears a resemblance to the constant-Q transform (CQT),
which is derived from the short-time Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q}$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the wavelet scalogram reaches the Heisenberg
theoretical limit of optimal time-frequency localization across the whole
frequency range, whereas the CQT has a fixed temporal resolution, constrained by the support of the STFT analyzing window.
Therefore, the wavelet scalogram has a better temporal localization at high
frequencies, which allows to observe steady amplitude modulations.

In all subsequent experiments, wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as a fourth-order Gammatone
wavelet with one vanishing moment \cite{Venkitaraman2014}.
